draft

### 2025-05-24 draft.tex
Updates from Overleaf

\subsection{The current state of bioimage analysis}
\label{sec:org8f5b33d}
The most widely used software for processing high-throughput biological images is CellProfiler \citep{stirlingCellProfiler4Improvements2021}. It can be used by experimental biologists with scant programming experience, for whom it has been a boon. By contrast, for computer-savvy scientists, its graphical user interface and lack of programmatic access means friction in pipelines comprised of multiple tools. CellProfiler is ideal for creating and iteratively adapting manually-defined workflows; for low-customization high-throughput analyses such as extracting features for image-based profiles, however, certain aspects become a hindrance. For example, coupling CellProfiler with other is a struggle, as building plug-ins to add functionality is a time and effort-consuming challenge that requires an understanding of its interfaces. The usage of CellProfiler also precludes the development of custom-made image-processing steps, such as applying an image filter before feature extraction. Lastly, CellProfiler depends on many Python and Java packages, increasing the likelihood of dependency conflicts. Containers mitigate this problem, but not without increased complexity and caveats.

@@

Already, cp\_measure  has been used in multiple experiments, including in non-biological contexts such as environmental monitoring \citep{ideharaExploringNileRed2025}. We therefore foresee cp\_measure being useful for a broad scientific community beyond image-based profiling.
\section{Future work}
\label{sec:org5cdbb12}
We propose in future work to contribute cp\_measure back to CellProfiler, which offers several benefits. It would ensure that the results from pipelines built with either tool will always be comparable, while also providing the opportunity to formalize the inputs and outputs of all measurements. It would also mean that any changes made in the leaner and more approachable cp\_measure propagate to CellProfiler, benefiting the community for which it was originally developed.

>>>>

\subsection{The current state of bioimage analysis}
\label{sec:org8f5b33d}
The most widely used software for processing high-throughput biological images is CellProfiler \citep{stirlingCellProfiler4Improvements2021}. It can be used by experimental biologists with scant programming experience, for whom it has been a boon. By contrast, for computer-savvy scientists, its graphical user interface and lack of programmatic access means friction in pipelines comprised of multiple tools. CellProfiler is ideal for creating and iteratively adapting manually-defined workflows; for low-customization high-throughput analyses such as extracting features for image-based profiles, however, certain aspects become a hindrance. For example, mix and matching CellProfiler with other image analysis tools is a struggle, as building plug-ins to add functionality is a time and effort-consuming challenge that requires an understanding of its interfaces. The usage of CellProfiler also precludes the development of custom-made image-processing steps, such as applying an image filter before feature extraction. Lastly, CellProfiler depends on many Python and Java packages, increasing the likelihood of dependency conflicts. Containers mitigate this problem, but not without increased complexity and caveats.

@@

Already, Cellprofiler features have been used in multiple experiments, including in non-biological contexts such as environmental monitoring \citep{ideharaExploringNileRed2025}. By association, we foresee cp\_measure being useful for a broad scientific community beyond image-based profiling.
\section{Future work}
\label{sec:org5cdbb12}
We propose in future work to contribute cp\_measure back to CellProfiler, which offers several benefits. It would ensure that the results from pipelines built with either tool will always be comparable, while also providing the opportunity to formalize the inputs and outputs of all measurements. It would also mean that any changes made in the leaner and more approachable cp\_measure propagate to CellProfiler, benefiting the community for which it was originally developed.
